Experiences designing and building a PRAGMA Cloud Scheduler

### 2015-08-14 paper.tex
Filled in pilot section with some text

\section{Introduction}

@@

\textbf{Pragma\_boot:}  rather than requiring a single cloud system at all sites, PRAGMA encourages the sites to  deploy any virtualization technology that best fits their expertise. However, enabling a virtual cluster to be ported to different cloud hosting environment is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.  In 2013, three pilot sites (UCSD, AIST, NCHC) developed a script for an automated virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.  Based on those experiences, the porting script was redesigned and reimplemented as \textit{pragma\_boot} toolkit with "drivers" to support  virtual cluster migration to virtual environments enabled on Rocks and OpenNebula clusters.

@@

NOTE THE FOLLOWING IS A PART OF AN IMPLEMENTATION NOT A REQUIREMENT, MOVE TO DIFFERENT SECTION THAT DESCRIBES IMPLEMENTATION: Sites  just need to run a SSH service and install the pragma\_boot package, which is a small Python package and can be installed via RPM and can be configured in less than an hour.

@@

Discuss creation of Cloud Scheduler Pilot.

>>>>

\section{Introduction}
\label{Sec:intro}

@@

\textbf{Pragma\_boot:}  rather than requiring a single cloud system at all sites, PRAGMA encourages the sites to  deploy any virtualization technology that best fits their expertise. However, enabling a virtual cluster to be ported to different cloud hosting environment is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.  In 2013, three pilot sites (UCSD, AIST, NCHC) developed a script for an automated virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.  Based on those experiences, the porting script was redesigned and reimplemented as \textit{pragma\_boot} toolkit with "drivers" to support  virtual cluster migration to virtual environments enabled on Rocks and OpenNebula clusters.  In 2014, a feature was added to pragma\_boot to download and boot virtual cluster images from Amazon CloudFront~\cite{cloudfront}.

@@

@@

One key requirement for our cloud scheduler was to enable users to start and manage virtual clusters on the PRAGMA cloud testbed with minimal effort. 
The basic workflow of how a user would start a virtual cluster and the steps are summarized below:

\begin{enumerate}
\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (i.e., IPOP, ViNE, or PRAGMA-ENT) that is needed.  
\item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  
\item The user selects the subset of resources they want, the timeframe, virtual cluster image, and configuration details and submits it to the Cloud scheduler.  Once the reservation is verified and confirmed, the user will get an email confirmation.  
\item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when it's ready for them to login.   
\end{enumerate}

Figure~\ref{Fig:Flow} shows the above steps in a use case sequence diagram where \textit{PRAGMA Booking} refers to our customized Booked instance.  PCC will monitor the health of the virtual cluster and when the reservation is close to expiring, will notify the user in case they want to extend their reservation.  Otherwise, when the time expires or the user chooses to end their reservation, the virtual cluster will be shut down.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\columnwidth]{figures/flow.pdf}
\caption{The flow of a user request to start a virtual cluster and the components that are involved in the cloud scheduler.}
\label{Fig:Flow}
\end{center}
\end{figure}

The other key requirement for the cloud scheduler was that it should require minimal effort for a PRAGMA sites to participate in the PRAGMA testbed.  Therefore, we opted for a simple client-server architecture, where the client is a small PRAGMA package that interfaces to a site's local cloud system.  Currently, the PRAGMA Package is just a SSH service and the pragma\_boot package, which is a small Python package that can be installed via RPM and can be configured in less than an hour.  The server component is composed of the PRAGMA Booking interface and the PCC which gets launched for per user.  The architecture for the PRAGMA cloud scheduler is shown in Figure~\ref{Fig:Arch}.

In our pilot implementation of the cloud scheduler, we made a number of temporary assumptions to simplify the design.

\begin{itemize}
\item Only one reservation can be made per resource.  While we made several changes in our pilot to customize the look of Booked, modifying it to allow for multiple reservations per resource would have required deeper changes that we wanted to defer to a later version.   
\item Virtual cluster images are already available at each site.  For our pilot, we worked with a small number of virtual cluster images and pre-installed them to each cloud system so they were readily accessible by pragma\_boot.  We had the following three images: 1) a basic SGE Rocks virtual cluster, 2) an AutoDoc virtual cluster, and 3) a Lifemapper compute virtual cluster.   
\item Since PCC is also under active development and was not easy to install, we used a simple stub to substitute for its functionality.  The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and looked for reservations that are ready to be activated.  Once it finds a reservation that need sto be activated, it launched pragma\_boot via SSH.  It does not  monitor the virtual cluster nor email the user when their reservation is about to expire.  
\item Only single site virtual clusters can be launched.  So far, we've only experimented with multi-site virtual clusters using IPOP and this requires the virtual cluster images to be pre-installed with IPOP and an automated configure script that will assign private IP addresses to each node on boot.  For Rocks virtual clusters, there is a Rocks roll~\cite{ipoproll} that will automate the IPOP installation and install the configure scripts.  
\end{itemize}

cloud scheduler would leverage 
To create a lightweight scheduler that enabled users to start and manage virtual clusters on leveraged the pragma\_boot, PCC, and software-defined networking tools that had been developed as part of PRAGMA 
As 
Our 
To schedule virtual clusters on the PRAGMA cloud testbed, 
Discuss creation of Cloud Scheduler Pilot.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\columnwidth]{figures/arch.pdf}
\caption{Architecture for the PRAGMA cloud scheduler.}
\label{Fig:Arch}
\end{center}
\end{figure}
